Centralized Programming for Distributed Control Planes

### 2016-01-27 paper/introduction.tex
intro rewrites

The centralized control planes of SDN, however, are not a panacea.
%
They require careful design and engineering to be robust to failures at scale---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure. Even ignoring failures, it is necessary for controller systems to
scale to meet the demands of larger, geographically distributed networks, 
and to react sufficiently quickly
to a changing environment. To address this challenge, some researchers are beginning to 
explore
multi-controller networks, with interacting controllers, thus bringing back distributed
control planes~\cite{mccauley2013extending,onos} and their current programming difficulties.  
%However, academic
%language design and implementation efforts have not kept pace.  For instance, work on many
%experimental SDN languages~\cite{frenetic,flowlog,vericon,merlin,netkat,kinetic,pga} has not yet shown how to implement fault tolerant
%multi-controller systems that support their high-level abstractions efficiently.
A second problem with existing SDN frameworks, and the languages used to
program them, is that while they provide many 
effective \emph{intra}-domain routing
abstractions, they fail to provide a \emph{coherent} means to enable 
\emph{inter}-domain communication of available routes.

Hence, in this paper, we have two central goals:
\begin{enumerate}
\item Design a new, high-level, declarative language with abstractions 
suitable for simultaneously expressing intra-domain routing, inter-domain 
routing and alternatives in case of failures.
\item Define algorithms for compiling these specifications in to 
configurations for off-the-shelf devices running standard
distributed control plane algorithms, while ensuring correct behavior
independent of the number of faults.
\end{enumerate}

To achieve the first goal, we borrowed heavily from the design of
recent high-level SDN languages based around specification of 
paths via regular expressions such as FatTire~\cite{fattire}, 
Merlin~\cite{foster:merlin}, and
NetKAT~\cite{netkat}.  However, one of the key differences is that
we incorporate expression of user preferences in to the algebra
of regular expressions and provide a semantics to our language
in terms of sets of ranked paths.  The preferences indicate possible
fail-over routes:  If too many faults occur, the network may be
disconnected, but as long as any path the user has expressed is
still available, the system guarantees the highest available ranked
path will be taken.  We also support network operators by defining a
set of common abbreviations for entering, leaving, and traversing
the user network.  Operators use familiar logical operations
to build routes in a modular and compositional manner from these 
primitives.  All abbreviations are translated in to our regular-expression-based
intermediate representation and compiled from there.

To achieve the second goal, we define program analysis and compilation
algorithms that translate the regular policies in to a graph-based
intermediate language and from there in to BGP.  During the
compilation process, our compiler automatically synthesizes per-device
import and export filters, local preferences, MED attributes, and 
community tags.  We chose BGP for pragmatic reasons:  it is an industry
standard and many data centers use it internally as well as externally.
Indeed, despite the presence of SDN, many networks will continue to
use BGP for the foreseeable future due to investments they have already
made in the infrastructure, the difficulty of transitioning to SDN,
and the scalability and fault-tolerance advantages of a distributed
control plane.  However, using BGP as a back end is not without its 
challenges: not all high-level specifications can be implemented correctly in
the presence of arbitrary failures.  Consequently, we define new
program analysis algorithms to detect specifications that cannot be implemented
and report our findings to the user, so he or she may fix the specification
early rather than waiting until configurations have been deployed.  
These algorithms differ from
standard BGP analysis algorithms~\cite{feamster:rcc,batfish} because they operate on representations of
our high-level specifications (which has the pleasant effect of simplifying our 
task) rather than directly on arbitrary BGP configurations.

To evaluate language and system, we have experimented with specifying policy
for several realistic data center and backbone networks. We have also
measured the performance of our compiler, demonstrating that our compilation
algorithms scale to topologies with several thousand routers.

>>>>

The centralized control planes of SDN, however, are not a panacea.
First, while many SDN systems provide effective \emph{intra}-domain routing
abstractions, allowing users to specify paths within their own network,
they fail to provide a coherent means to enable specification of preferred
\emph{inter}-domain routes.  Second, centralized control planes
require careful design and engineering to be robust to failures at scale---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure. Even ignoring failures, it is necessary for controller systems to
scale to meet the demands of larger, geographically distributed networks, 
and to react sufficiently quickly
to a changing environment. To address this challenge, some researchers are beginning to 
explore
multi-controller networks, with interacting controllers, thus bringing back distributed
control planes~\cite{mccauley2013extending,onos} and their current programming difficulties.  
%However, academic
%language design and implementation efforts have not kept pace.  For instance, work on many
%experimental SDN languages~\cite{frenetic,flowlog,vericon,merlin,netkat,kinetic,pga} has not yet shown how to implement fault tolerant
%multi-controller systems that support their high-level abstractions efficiently.

Hence, in this paper, we have two central goals:
\begin{enumerate}
\item Design a new, high-leve language with abstractions 
for expressing intra-domain routing, inter-domain 
routing and routing alternatives in case of failures.
\item Define algorithms for compiling these specifications in to 
configurations for devices running standard
distributed control plane algorithms, while ensuring correct behavior
independent of the number of faults.
\end{enumerate}

To achieve the first goal, we borrowed heavily from the design of
recent high-level SDN languages such as FatTire~\cite{fattire}, 
Merlin~\cite{foster:merlin}, and
NetKAT~\cite{netkat} by using regular expressions to specify paths
through networks.  However, our design also contains several key
differences from existing languages.  The most important difference is semantic:  the paths specified
by programmers can extend from outside the operator's network to inside
the network, across several devices internally, and then out again.  This design
choice allows users to specify preferences about both external and internal 
routes in the same uniform notation.  No SDN programming language provides
similar facilities.  In addition,
we incorporate expression of user preferences in to the algebra
of regular expressions and provide a semantics to our language
in terms of sets of ranked paths.  These preferences indicate possible
fail-over routes:  If too many faults occur, the network may be
disconnected, but as long as any path the user has expressed is
still available, the system guarantees the highest-ranked, available
path will be taken.  We also support network operators by defining a
set of common abbreviations for entering, leaving, and traversing
the user network and we provide abstractions for managing common features of
control plane algorithms such as route aggregation.  Operators use familiar logical operations
to build routes in a modular and compositional manner from these 
primitives.  Our abbreviations are translated in to our regular-expression-based
intermediate representation and compiled from there.

To achieve the second goal, we define program analysis and compilation
algorithms that translate the regular policies in to a graph-based
intermediate language and from there in to BGP.  During the
compilation process, our compiler automatically synthesizes per-device
import and export filters, local preferences, MED attributes, and 
community tags.  We chose BGP for pragmatic reasons:  it is an industry
standard and many data centers use it internally as well as externally.
Indeed, despite the presence of SDN, many networks will continue to
use BGP for the foreseeable future due to investments they have already
made in the infrastructure, the difficulty of transitioning to SDN,
and the scalability and fault-tolerance advantages of a distributed
control plane.  However, using BGP as a back end is not without its 
challenges: not all high-level specifications can be implemented correctly in
the presence of arbitrary failures.  Consequently, we define new
program analysis algorithms to detect specifications that cannot be implemented
and report our findings to the user, so he or she may fix the specification
early rather than waiting until configurations have been deployed.  We also
analyze route aggregation to detect the possibility of failure-induced black-holes.
These algorithms differ from
standard BGP analysis algorithms~\cite{feamster:rcc,batfish} because they operate on 
our graph-based intermediate representation, which operates at a higher level
of abstraction than raw BGP configurations (and which has the pleasant effect of simplifying our 
task).

To evaluate the \sysname{} language and system, we have experimented with specifying policy
for several realistic data center and backbone networks and compared these specifications
with existing operating specifications. We have also
measured the performance of our compiler, demonstrating that our compilation
algorithms scale to topologies with several thousand routers.
